---
title: "R Notebook"
output: html_notebook
---

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```


Pull in data
```{r}
Trees <- read.csv(here("Data", "Forest_Trees_17_21.csv"))
Trees_regen <- read.csv(here("Data", "Forest_Trees_and_Regen_2021.csv"))
Regen <- read.csv(here("Data", "Forest_Regeneration_17_21.csv"))
```


##Part 1: Find out how many forests each species is present at

Make uniform species titles
```{r}
Trees$Common.name <- factor(Trees$Common.name)
levels(Trees$Common.name) #there are 42 levels, so 42 "species" with many repeats


#time to fix the repeats!
wrong <- which(Trees$Common.name == "Black cherry"| Trees$Common.name == "Black cherry") 

Trees$Common.name[wrong]<-"black cherry" #rename the wrong black cherries
Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 41 levels 
```

```{r}
#muscle wood is the next species I see in testing with duplicates. I will fix it

bad <- which(Trees$Common.name == "muscle wood"| Trees$Common.name == "muscle wood") 
Trees$Common.name[bad]<-"musclewood" 

no <- which(Trees$Common.name == "Musclewood"| Trees$Common.name == "Musclewood") 
Trees$Common.name[no]<-"musclewood" 


Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 39 levels 
```

```{r}
#time to fix quaking aspen

horrible <- which(Trees$Common.name == "Quaking aspen"| Trees$Common.name == "Quaking aspen")
Trees$Common.name[horrible]<-"quaking aspen" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 38 levels 
```

```{r}
#fix the red maples

terrible <- which(Trees$Common.name == "Red maple"| Trees$Common.name == "Red maple")
Trees$Common.name[terrible]<-"red maple" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 37 levels 
```

```{r}
#fix the red oaks
not <- which(Trees$Common.name == "Red oak"| Trees$Common.name == "Red oak")
Trees$Common.name[not]<-"red oak" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 36 levels 
```

```{r}
#fix the red pine
ugh <- which(Trees$Common.name == "Red pine"| Trees$Common.name == "Red pine")
Trees$Common.name[ugh]<-"red pine" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 35 levels 
```

```{r}
#fix the red spruce
meh <- which(Trees$Common.name == "Red spruce"| Trees$Common.name == "Red spruce")
Trees$Common.name[meh]<-"red spruce" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 34 levels 
```

```{r}
#fix the hickory
why <- which(Trees$Common.name == "Shagbark hickory"| Trees$Common.name == "Shagbark hickory")
Trees$Common.name[why]<-"shagbark hickory" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 33 levels 
```

```{r}
#fix the sugar maple
ok <- which(Trees$Common.name == "Sugar maple"| Trees$Common.name == "Sugar maple")
Trees$Common.name[ok]<-"sugar maple" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 32 levels
```

```{r}
#fix the white ash
right <- which(Trees$Common.name == "White ash"| Trees$Common.name == "White ash")
Trees$Common.name[right]<-"white ash" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 31 levels
```

```{r}
#fix the white oak
huh <- which(Trees$Common.name == "White oak"| Trees$Common.name == "White oak")
Trees$Common.name[huh]<-"white oak" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 30 levels
```

```{r}
#fix the white pine
fix <- which(Trees$Common.name == "White pine"| Trees$Common.name == "White pine")
Trees$Common.name[fix]<-"white pine" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 29 levels
```

```{r}
#fix the yellow birch
last <- which(Trees$Common.name == "Yellow birch"| Trees$Common.name == "Yellow birch")
Trees$Common.name[last]<-"yellow birch" 

Trees$Common.name<-factor(Trees$Common.name)
levels(Trees$Common.name) #now there are only 28 levels

#I think those were all the duplicates
```


Now, with the cleaned data, find how many forests each species is present at and the number of species at each plot
```{r}
 Tree_Rich <- Trees %>% group_by(Common.name, Forest, PlotNum) %>%
  summarize(Presence = length((Forest)))

 #I can see that red maple is the only species present at all 5 forests
```


```{r}
Tree_Spp <- Trees %>%
  group_by(Common.name, Forest) %>%
  summarize(Obs = length(Common.name)) 
```

```{r}
Tree_pres <- Trees %>%
  group_by(Common.name) %>%
  summarize(Num = length(unique(Forest)))
```




##Part 2: find the number of each species at each site 

set up 
```{r}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(here)
```


pull in data
```{r}
DF <- read.csv(here("Data", "Forest_Trees_and_Regen_2021.csv"))
```

clean species
```{r}
unique(DF$sci_name)
length(unique(DF$sci_name)) #says there are 31 species

#can see I need to remove NA and blank values  
DF <- DF[-which(is.na(DF$sci_name)), ]
DF <- DF[-which(DF$sci_name == ""), ]

#check to see if it worked. I don't see any NAs or blanks!
unique(DF$sci_name)
length(unique(DF$sci_name)) #i now have 29, so I know two were deleted
```


now make a dataframe that displays how many of each species is found in each forest
```{r}
Num_obs <- DF %>% group_by(sci_name, Forest) %>%
  summarize(num_trees = length(sci_name))
```

make a dataframe that shows how many of each species is found in each plot
```{r}
Num_obs_site <- DF %>% group_by(sci_name, Forest, Plot) %>%
  summarize(num_trees = length(sci_name))
```

